---
title: "Exploring Boston Dataset"
author: "Roxana Butenco"
date: "7/2/2019"
output:
  html_document: 
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Descriptive Statistics
In this post we'll look at Boston dataset in more detail and will seek significant predictors in relation to crime. 
Will look at the correlation between all variables to analyse each correlation coefficient, getting a first impression of the relationship between variables.
At the moment the goal is to find what kind of regression modelling, if any, can be applied to data for better crime prediction (for example if linear trends are visible, we may fit a linear regression model).

## The Dataset
We are going to use the Boston dataset in the MASS package. This dataset contains information collected by the U.S Census Service concerning housing in the area of Boston Mass. The dataset is small in size with only 506 cases. Our goal is to predict crime using all other variables available so crim is our variable of interest and all other are the predictors.

## Data Ingestion and Exploration
Specify the working directory and read in data from a csv file as we have already downloaded our dataset:
```{r}
setwd("~/RDemo/BostonRegression")
Boston = read.csv(file="Boston.csv",header=TRUE,sep=",")
View(Boston) #invoke the data viewer that also contain some basic exploratory features
fix(Boston) #open a quick data editor
??Boston #get familliar with the 14 attributes and their description from our dataset original source
```
Data accuracy is important so first we need to check for missing values:
```{r}
colSums(is.na(Boston))
```
None. We can proceed to some descriptive statistics:
```{r}
str(Boston)
cat("The mean : ", mean(Boston$crim))
cat("The median : ", median(Boston$crim))
cat("The mode : ", mode(Boston$crim))
cat("The standard deviation : ", sd(Boston$crim), "\n")
```
## Correlation
Correlation matrix visualization with Pearson correlation coefficient.
Predictor variables correlation chart 
```{r message=FALSE, warning=FALSE, out.width="80%", fig.align='center'}
library(zoo)
library(xts)
library(PerformanceAnalytics)
chart.Correlation(Boston, histogram=TRUE, pch=19)
```
## Data Distribution
Let's look at how data is distributed (predictor variables versus target variable crim).  
Scatterplots of target variable crim with all predictor variables less chas:
```{r warning=FALSE, out.width="80%", fig.align='center'}
library(tidyverse)
Boston %>%
  gather(-crim, key = "var", value = "value") %>%
  filter(var != "chas") %>%
  ggplot(aes(x = value, y = crim)) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~ var, scales = "free") +
  theme_bw()
```
Predictors histograms less chas:
```{r warning=FALSE, out.width = "80%", fig.align='center'}
Boston %>% 
  gather(-crim, key = "var", value = "value") %>%
  filter(var != "chas") %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~ var, scales = "free") +
  theme_bw()
```
A closer look at correlation between crim and zn: 
```{r warning=FALSE, out.width = "60%", fig.align='center'}
my_data <- Boston[, c(2,1)]
chart.Correlation(my_data, histogram=TRUE, pch=19)
```
## Outliers
By looking at boxplots we can easily see if there are any outliers:
```{r warning=FALSE, out.width = "80%", fig.align='center'}
Boston %>% #Predictors boxplots -chas where we can clearly see the outliers 
  gather(-crim, key = "var", value = "value") %>%
  filter(var != "chas") %>%
  ggplot(aes(x = '',y = value)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 1) +
  facet_wrap(~ var, scales = "free") +
  theme_grey()
```
Is the correlation between zone and crim statistically significant:
```{r warning=FALSE, out.width = "60%", fig.align='center'}
ggplot(data=Boston)+
   geom_point(mapping = aes(x=zn, y=crim)) +
    theme_bw()
cor.test(Boston$zn,Boston$crim)#
```
By looking at the above plot and correlation coefficient, it doesn't seem there is much correlation between zn and crim, 
or at least it's not statistically significant.

## Conclusion
Our response variable crim has a moderate uphill relationship with rad, tax, lstat (e.g. with a unit change in rad, 
crim increases), a week positive relationship with nox, indus and age and it has a week negative relation with medv, black and dis 
(crim increases as medv decreases). 
Crim shows almost no linear relationship with ptratio, rm and zn and the correlation doesn't mean much in these cases.
So, the only safe conclusion here is that a linear trend may exist between rad, tax, lstat, nox, indus, age, medv, black, dis and crim.